Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-10-03 arxiv/examples.tex
revise

\section{Illustrative Examples}\label{sec_example}

%We provide precise explanations to the phenomena of negative transfer in multi-task learning.
%We provide tight bounds on the bias and variance of the multi-task estimator for two tasks.
%The results in Section \ref{sec_general} show when multi-task learning outperforms single-task learning for general task covariance matrices and ground truth model parameters.
%In particular, the general conditions depend on specific properties of task data such as the covariate shift matrix.
%However, the conditions given by the limiting bias and variance of multi-task learning can be difficult to interpret.
%The goal of this section is to address this issue.
%We interpret the limiting bias and variance equations in an isotropic covariance setting that is described below and a simplified covariate-shifted setting.
%We show that three properties of task data affect the performance of multi-task learning as follows.
%(i) \textit{Task similarity}: We explain the phenomenon of negative transfer precisely as tasks become more different.
%(ii) \textit{Sample ratio}: We further explain a curious phenomenon where increasing the sample ratio between the source and target task helps initially, but hurts eventually.
%(iii) \textit{Covariate shift}: Finally, we show that as the sample size of the source task increases, we show that the covariate shift worsens the performance of the multi-task estimator.
%Finally, we provide a extend our results from two tasks to many tasks with the same features.
%We explain from three perspectives, including \textit{task similarity}, \textit{sample size} and \textit{covariate shift}.
%We show how negative transfer occurs by varying task similarity or sample size.
%Then we show that when source task sample size becomes large, covariate shift causes more negative effects.
We discuss several examples to illustrate our results.
Through these examples, we provide generalization bounds of hard parameter sharing estimators that depend explicitly on the sample sizes and the covariance matrices of the tasks.
We use our generalization bounds to interpret several empirical phenomena that arises from applying hard parameter sharing in practice.
For our examples, we assume that a random-effects model holds.
We assume every $\beta_i$ consists of two random components, one that is shared among all tasks and one that is task-specific.
Thus, each task contributes a certain amount to the shared component and injects a task-specific bias.
%	The labels are $Y_i = X_i\beta_i + \varepsilon_i$, where $\e_i$ consists of i.i.d. entries with mean zero and variance $\sigma^2$.
%	For our purpose, it is enough to think of the order of $d$ being $1/\sqrt{p}$ and $pd^2/\sigma^2$ being constant.
%The more precise conditions on the relations between $d^2$, $\sigma^2$ and $\kappa^2$ are given in  \eqref{choiceofpara}.
%	We assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in Appendix \ref{sec_maintools}.

@@

\begin{corollary}[Derived from Theorem \ref{thm_main_RMT}]\label{cor_MTL_loss}
	In the setting described above, assume further that a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that 
	the signal and noise variances satisfy 
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	\be\label{choiceofpara0}
	\sigma^2 \lesssim  \kappa^2   ,\quad  d^2 \ll {\kappa^2}.
	\ee
	%Let $B, A_1, A_2$ be the global minimum of equation \eqref{eq_mtl}. Then
	The prediction loss of the hard parameter sharing estimator for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le 
	  \left|L(\hat{\beta}_2^{\MTL}) -\frac{\sigma^2}{\rho_1+\rho_2-1} - pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\right|  \le & \left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	& +Cp \left[p^{-c_\varphi}(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + d^4/\kappa^2\right],\nonumber
	 \end{align}
	 for a large constant $C>0$.
\end{corollary}

>>>>

%\section{Illustrative Examples}\label{sec_example}

%We provide precise explanations to the phenomena of negative transfer in multi-task learning.
%We provide tight bounds on the bias and variance of the multi-task estimator for two tasks.
%The results in Section \ref{sec_general} show when multi-task learning outperforms single-task learning for general task covariance matrices and ground truth model parameters.
%In particular, the general conditions depend on specific properties of task data such as the covariate shift matrix.
%However, the conditions given by the limiting bias and variance of multi-task learning can be difficult to interpret.
%The goal of this section is to address this issue.
%We interpret the limiting bias and variance equations in an isotropic covariance setting that is described below and a simplified covariate-shifted setting.
%We show that three properties of task data affect the performance of multi-task learning as follows.
%(i) \textit{Task similarity}: We explain the phenomenon of negative transfer precisely as tasks become more different.
%(ii) \textit{Sample ratio}: We further explain a curious phenomenon where increasing the sample ratio between the source and target task helps initially, but hurts eventually.
%(iii) \textit{Covariate shift}: Finally, we show that as the sample size of the source task increases, we show that the covariate shift worsens the performance of the multi-task estimator.
%Finally, we provide a extend our results from two tasks to many tasks with the same features.
%We explain from three perspectives, including \textit{task similarity}, \textit{sample size} and \textit{covariate shift}.
%We show how negative transfer occurs by varying task similarity or sample size.
%Then we show that when source task sample size becomes large, covariate shift causes more negative effects.
%We discuss several examples to illustrate our results.
%Through these examples, we provide generalization bounds of hard parameter sharing estimators that depend explicitly on the sample sizes and the covariance matrices of the tasks.
%We use our generalization bounds to interpret several empirical phenomena that arises from applying hard parameter sharing in practice.
%For our examples, we assume that a random-effects model holds.
%We assume every $\beta_i$ consists of two random components, one that is shared among all tasks and one that is task-specific.
%Thus, each task contributes a certain amount to the shared component and injects a task-specific bias.
%	The labels are $Y_i = X_i\beta_i + \varepsilon_i$, where $\e_i$ consists of i.i.d. entries with mean zero and variance $\sigma^2$.
%	For our purpose, it is enough to think of the order of $d$ being $1/\sqrt{p}$ and $pd^2/\sigma^2$ being constant.
%The more precise conditions on the relations between $d^2$, $\sigma^2$ and $\kappa^2$ are given in  \eqref{choiceofpara}.
%	We assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in Appendix \ref{sec_maintools}.

@@

\begin{corollary}[Derived from Theorem \ref{thm_main_RMT}]\label{cor_MTL_loss}
	In the setting described above, assume further that a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that 
	the signal and noise variances satisfy 
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	\be\label{choiceofpara0}
	\sigma^2 \lesssim  \kappa^2   ,\quad  d^2 \ll {\kappa^2}.
	\ee
	%Let $B, A_1, A_2$ be the global minimum of equation \eqref{eq_mtl}. Then
	The prediction loss of the hard parameter sharing estimator for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le 
	  \left|L(\hat{\beta}_2^{\MTL}) -\frac{\sigma^2}{\rho_1+\rho_2-1} - d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\right|  & \le  \left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + \frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
	 for a large constant $C>0$.
\end{corollary}
